Probability Aggregation in Time-Series: Dynamic Hierarchical Modeling of Sparse Expert Beliefs

### 2013-09-24 paper.tex
Final edits

This paper adds another layer of complexity by assuming a heterogeneous set of experts, most of whom only make one or two probability assessments over the hundred or so days before the event resolves. This means that the decision-maker faces a different group of experts  every day, with only a few experts returning later on for a second round of advice. The problem at hand is therefore strikingly different from many time-series estimation problems, where one has an observation at every time point -- or almost every time point. As a result, standard time-series procedures like ARIMA (see, e.g., \citet{mills1991time}) are not directly applicable. This paper introduces a time-series model that incorporates self-reported expertise and captures a sharp and well-calibrated estimate of the crowd belief. The model is highly interpretable and can be used for:
\begin{itemize}
\item borrowing strength across the hierarchy to analyze the under- and overconfidence in different groups of experts,
%\item confidence bands for the crowd belief,
\item accurate probability forecasts, and
\item many question-specific quantities that have easy interpretations, such as expert disagreement and problem difficulty, and can be used to gain novel insight in the social sciences. 
\end{itemize}

This paper begins by describing our geopolitical database. The paper then introduces a dynamic hierarchical model for capturing the crowd belief. The model is estimated in a two-step procedure: first, a sampling step produces constrained parameter estimates via Gibbs sampling (see \citet{geman1984stochastic} for the original introduction of Gibbs sampling); second, a calibration step transforms these estimates to their unconstrained equivalents via a one-dimension optimization procedure. An extension of this model to polychotomous outcomes is briefly discussed before model evaluation. The first evaluation section uses synthetic data to study how accurately the two-step procedure can estimate parameter values. The second evaluation section applies the model to real-world forecasting data. The paper concludes with a discussion on future research directions and model limitations.

@@

The experts made updates on a very infrequent basis: the average number of forecasts per expert was around 0.017 forecasts per day, and the average group-level response rate was around 13.5 forecasts per day. Given that the group of experts is large and diverse, the resulting dataset is very sparse. Tables \ref{DataStats} and \ref{ExpertiseTable} provide more relevant summary statistics on the data. Notice that the distribution of the self-reported expertise is skewed to the right and that some questions remained active longer than others.  For more details on the dataset and its collection see \citet{ungar2012good}.

@@

\subsection{Sampling Step}
\label{sampling_step}
Given that $\left(a \boldsymbol{b}, X_{t,k}/a, a^2 \tau_k^2\right) \neq \left(\boldsymbol{b}, X_{t,k}, \tau_k^2\right)$ for any $a > 0$ yield the same  likelihood for $\boldsymbol{Y}_{t,k}$, the model as described by Equations (\ref{observedpr}) and (\ref{hiddenpr}) is not identifiable. As a result, the parameter estimates tend to drift during the sampling process. A well-known solution is to choose one of the elements of $\boldsymbol{b}$, say $b_3$, as the reference point and fix $b_3 = 1$.  Denote the constrained version of the model by
\begin{eqnarray*}
\boldsymbol{Y}_{t, k} &=&  \boldsymbol{M}_k \boldsymbol{b} (1) X_{t, k}(1)+ \boldsymbol{v}_{t, k}\\
X_{t, k}(1) &=& \gamma_k(1) X_{t-1, k}(1) + w_{t, k}\\
\boldsymbol{v}_{t, k} | \sigma^2_k(1) &\stackrel{i.i.d.}{\sim}& \mathcal{N}_{I_k}(\boldsymbol{0}, \sigma^2_k(1) \boldsymbol{I}_{I_k})\\
w_{t, k} | \tau^2_k(1) &\stackrel{i.i.d.}{\sim}& \mathcal{N}\left(0, \tau^2_k(1)\right),
\end{eqnarray*}
where the trailing input parameter emphasizes the constraint $b_{3} = 1$. Given that this version is identifiable, estimates of the model parameters can be obtained. Denote the estimates by placing a hat on the parameter symbol. For instance, $\hat{\boldsymbol{b}}(1)$ and $\hat{X}_{t, k}(1)$ represent the estimates of $\boldsymbol{b}(1)$ and $X_{t, k}(1)$, respectively. These estimates are found via Gibbs sampling that only makes use of standard distributions. See Appendix \ref{appendix} for the technical details of the sampling step, and, e.g., \citet{gelman2003bayesian} for a discussion on the general principles of Gibbs sampling.

\subsection{Calibration Step}
\label{calibration_step}
Given that the model parameters can be estimated by fixing $b_3$ to any constant, the next step is to search for the constant that gives an optimally sharp and calibrated estimate of the hidden process.  This section introduces an efficient procedure that finds the optimal constraint without requiring any additional runs of the sampling step. First, assume that parameter estimates  $\hat{\boldsymbol{b}}(1)$ and $\hat{X}_{t, k}(1)$ have already been obtained via the constrained sampling step described in Section \ref{sampling_step}. Given that for any $\beta \in \R / \{0\}$,
\begin{eqnarray*}
\boldsymbol{Y}_{t, k} &=&  \boldsymbol{M}_k \boldsymbol{b} (1) X_{t, k}(1)+ \boldsymbol{v}_{t, k}\\
 &=&   \boldsymbol{M}_k \left(\boldsymbol{b} (1) \beta \right) \left(X_{t, k}(1)/\beta\right)  + \boldsymbol{v}_{t, k}\\
&=&   \boldsymbol{M}_k \boldsymbol{b} (\beta) X_{t, k}(\beta) + \boldsymbol{v}_{t, k},
\end{eqnarray*}
the parameter values under $b_{3} = \beta$ can be obtained from $\boldsymbol{b} (\beta) = \boldsymbol{b} (1) \beta $ and $X_{t, k}(\beta) = X_{t, k}(1)/\beta$. 
%Furthermore, an estimate of $X_{t, k}(\beta)$ is given by $\hat{X}_{t, k}(1)/\beta$. 
This means that $X_{t,k} = X_{t,k}(1)/\beta$ when $\beta$ is equal to the true value of $b_3$. Given that the hidden process $X_{t, k}$ is assumed to be sharp and well-calibrated, $b_3$ can be estimated with the value of $\beta$ that simultaneously maximizes the sharpness and calibration of $\hat{X}_{t, k}(1) / \beta$. A natural criterion for this maximization is given by the class of \textit{proper scoring rules} that combine sharpness and calibration (\citet{gneiting2008rejoinder, buja2005loss}). Due to the possibility of \textit{complete separation} in any one question (see, e.g., \citet{gelman2008weakly}), the maximization must be performed over multiple questions. Therefore,
\begin{eqnarray}
 \hat{\beta}&=&  \argmax_{\beta \in \R / \{0\}} \sum_{k=1}^K \sum_{t=1}^{T_k}  S\left(Z_k, \hat{X}_{k,t}(1) / \beta \right) \label{OSE}
\end{eqnarray}
where $Z_k \in \{0, 1\}$ indicate whether the event associated with the $k$th question happened $\left(Z_k = 1\right)$ or did not happen $\left(Z_k = 0\right)$. The function $S$ is a strictly proper scoring rule such as the negative Brier score (\citet{Brier})
\begin{eqnarray*}
S_{BRI}(Z, X) &=& -(Z - \logit^{-1}(X))^2
\end{eqnarray*}
or the logarithmic score (\citet{good1952rational})
\begin{eqnarray*}
S_{LOG}(Z, X) &=& Z \log\left(\logit^{-1}(X)\right) + (1-Z) \log\left(1-\logit^{-1}(X)\right)
\end{eqnarray*}
Given that it is not clear which rule should be used for predicting geopolitical events, the \textit{Sample-And-Calibrate} procedure is evaluated  separately under both rules in Sections \ref{syntheticData} and \ref{realData}. Once $\hat{\beta}$ has been computed, estimates of the unconstrained model parameters are given by
\begin{eqnarray}
 \hat{X}_{t,k}&=& \hat{X}_{k,t}(1) / \hat{\beta} \nonumber\\
 \hat{\boldsymbol{b}}&=& \hat{\boldsymbol{b}}(1) \hat{\beta} \nonumber\\
   \hat{\tau}_{k}^2&=& \hat{\tau}_{k}^2(1)/ \hat{\beta}^2\nonumber\\
  \hat{\sigma}_{k}^2&=& \hat{\sigma}_{k}^2(1)\nonumber\\
  \hat{\gamma}_{k}&=& \hat{\gamma}_{k}(1)\nonumber
\end{eqnarray}
Notice that estimates of $\sigma^2_k$ and $\gamma_k$ are not affected by the constraint. Therefore their constrained and unconstrained versions are the same.

@@

If the future event can take upon $M > 2$ possible outcomes, the hidden state $X_{t,k}$ must extended to a vector of size $M-1$ and one of the outcomes, e.g., the $M$th one, must be chosen as the base-case to ensure that the probabilities will sum to one at any given time point. Each of the remaining  $M-1$ possible outcomes is represented by an observed process similar to Equation (\ref{observedpr}). Given that this multinomial extension is equivalent to having $M-1$ independent binary-outcome models, the estimation and properties of the model are easily extended to the multi-outcome case. This paper focuses on binary-outcomes because it is the simplest and most commonly encountered setting in practice.

@@

The goal in this section is to evaluate the ability of the SAC-procedure to capture true parameter values. The synthetic data are not generated directly from the model for two reasons: (i) showing good performance on a dataset directly generated from the model assumptions is hardly any news, and (ii) generating data from the dynamic model description does not produce well-calibrated hidden states. The latter is important for our interpretation of the hidden process as a sharp and well-calibrated version of the crowd belief.

To ensure that the hidden process for question $k$ is well-calibrated, generate a path of the standard Brownian motion until time $T_k$. If $Z_{t,k}$ denotes the value of the path at time $t$, then
\begin{eqnarray*}
Z_k &=& \mathbbm{1}\left( Z_{T_k, k}  > 0\right)\\
 X_{t,k} &=& \logit \left[ \Phi \left(\frac{Z_{t, k}}{\sqrt{T_k- t}} \right) \right]
\end{eqnarray*}
gives a sequence of $T_k$ calibrated logit-probabilities for the event $Z_k = 1$. Using this procedure we generate the hidden process for $K$ questions, each with a time horizon of $T_k = 101$. The questions involve 25 experts allocated evenly among five expertise groups. Each expert gives one probability forecast per day with the exception of time $t = 101$, when the event resolves. These are generated by applying bias and noise to the hidden process $X_{t,k}$ for $t = 1, \dots, 100$ (see Equation (\ref{observedpr})). In this fashion, the simulation iterates over a three-dimensional grid of parameter values:
\begin{eqnarray*}
\sigma^2 &\in& \{1/2, 1, 3/2, 2, 5/2\}\\
\beta &\in& \{1/2, 3/4, 1, 4/3, 2/1\}\\
K &\in& \{20, 40, 60, 80, 100\},
\end{eqnarray*}
where $\beta$ gives the bias vector by letting $\boldsymbol{b} = [1/2, 3/4, 1, 4/3, 2/1]^T \beta$. Every grid point is used  40 times to generate a synthetic dataset. The SAC-procedure is run on each dataset for 200 iterations of which the first 100 are used for burn-in. The accuracy of the estimated hidden process and bias vector are measured with the average quadratic loss in the probability space, $\sum_{k=1}^K \sum_{t=1}^{100} ( \logit^{-1}(\hat{X}_{t,k}) - \logit^{-1}(X_{t,k}))^2 / (100K)$, and the average quadratic loss, $|| \boldsymbol{b} - \hat{\boldsymbol{b}}||^2/5$, respectively.

@@

\begin{figure}[h!]
\vspace{-4.5em}
\centering
 \hspace{-31em}	\includegraphics{Figures/LegendSynthetic} % requires the graphicx package
\vspace{-1.5em}

@@

\vspace{-1em}

@@

\caption[Optional caption for list of figures]{Comparing SAC that optimizes over the logarithmic score ($\text{SAC}_{\text{LOG}}$) and SAC that optimizes over the Brier score ($\text{SAC}_{\text{BRI}}$) under synthetic data. The top row presents the accuracy to capture the hidden process. The bottom row presents the accuracy to capture the bias vector. }
\label{Synthetic}
\end{figure}

Figure \ref{Synthetic} summarizes the results in a series of plots. The first row presents the marginal effects of the three grid variables on the accuracy to which the SAC-procedure estimates the hidden process.  The second row is similar in nature but measures the accuracy in estimating the bias vector instead.  Given that the plots were computed by choosing one grid variable (e.g., $\beta$) at a time and averaging over the remaining two variables  (e.g., $K$ and $\sigma^2$), each value in the plots represents an average of a $40 \times 5^2 = 1,000$ values. Based on these results, both $\text{SAC}_{\text{LOG}}$ and  $\text{SAC}_{\text{BRI}}$ estimate the hidden process and the bias vector very accurately. Apart from $\text{SAC}_{\text{LOG}}$ outperforming $\text{SAC}_{\text{BRI}}$ uniformly across all parameter values, the two approaches behave very similarly.  They make great use of the increasing number of questions, suffer slightly from the increasing level of noise in the expert logit-probabilities, and present a trade-off between estimating $X_{t,k}$ and $\boldsymbol{b}$ under different values of $\beta$. This trade-off arises from \ville{WHY DOES THIS HAPPEN?}

@@

\section{Geopolitical Data Results}
\label{realData}
\noindent
This section presents results related to the real-world data described in Section \ref{data}. The goal is to provide application specific insight by discussing the specific research objectives itemized in Section \ref{intro} and also to evaluate the \textit{Sample-And-Calibrate}-procedure in terms of predictive power and calibration. Before presenting the results, however, we discuss two practical matters that must be taken into account when aggregating real-world  probability forecasts.

@@

The second matter is related to the distribution of the class labels in the data. If the set of occurrences is much larger than the set of non-occurrences (or \textit{vice versa}), the dataset is called \textit{imbalanced}. On such data the modeling procedure can end up over-focusing on the larger class, and as a result, give very accurate forecast performance over the larger class at the cost of performing poorly over the smaller class (see, e.g., \citet{chen2009learning, wallace2012class}). Fortunately, it is often possible to use a well-balanced version of the data. The first step is to find a partition $S_0$ and $S_1$ of the question indices $\{1, 2, \dots, K\}$ such that the equality $\sum_{k \in S_0} T_k  = \sum_{k \in S_1} T_k$ is as closely approximated as possible. This is equivalent to an NP-hard problem known in computer science as the \textit{Partition Problem}:  determine whether a given set of positive integers can be partitioned into two sets such that the sums of the two sets equal to each other (see, e.g., \citet{karmarkar1982differencing, hayes2002easiest}). A simple solution is to use a greedy algorithm that iterates through the values of $T_k$ in descending order, assigning each $T_k$ to the subset that currently has the smaller sum (see, e.g. \citet{kellerer2004knapsack, gent1996phase} for more details on the Partition Problem). After finding a well-balanced partition, the next step is to assign the class labels such that the labels for the questions in $S_x$ are equal to $x$ for $x = 0$ or $1$. Recall from section \ref{calibration_step} that $Z_k$ represents the event indicator associated with the $k$th question. To define a balanced set of indicators $ \tilde{Z}_k$ for all $k \in S_x$, let
\begin{align*}
 \tilde{Z}_k &= x\\
\tilde{p}_{i,t,k} &=  \begin{cases} 
1-p_{i,t,k}, & \text{if } Z_k = 1-x,\\
p_{i,t,k}, & \text{if } Z_k = x,
\end{cases}
\end{align*}
where $i = 1, \dots, I_k$, and $t = 1, \dots, T_k$. The resulting set $$\left\{\left(\tilde{Z}_k, \{\tilde{p}_{i,t,k} | i = 1, \dots, I_k, t = 1, \dots, T_k \}\right) \right\}_{k=1}^K$$ is a balanced version of the data. 
This procedure was used to balance our real-world dataset both in terms of events and time points. The final output splits the events exactly in half ($|S_0| = |S_1| = 83$) such that number of time points in the first and second halves are 8,737 and 8,738, respectively.

\subsection{Out-of-Sample Forecasting}
\label{forecasting}
This section is motivated by decision-making. 
%For instance, recall our example with an investor who consults a group of experts on a daily basis in order to assess whether a stock index will end up trading above some threshold on a given date. 
The goal is to evaluate the out-of-sample predictive performance of \textit{Sample-And-Calibrate} against several other probability aggregation procedures. The models are allowed to utilize a training set before making predictions on an independent testing set. To clarify some of the upcoming notation, let $S_{train}$ and $S_{test}$ be index sets that partition the data into training and testing sets of sizes $|S_{train}| = N_{train}$ and $|S_{test}| = 166 - N_{test}$, respectively. This means that the $k$th question is in the training set if and only if $k \in S_{train}$. The competing models are as follows:

@@

\item \textit{The Sample-And-Calibrate procedure both under the Brier ($\text{SAC}_{\text{BRI}}$) and the logarithmic score ($\text{SAC}_{\text{LOG}}$)}. The model is first fit on the training set by running the sampling step for 3,000 iterations of which the first 500 iterations are used for burn-in. After thinning by only keeping every fifth observation, the calibration step is performed for the remaining 500 observations. The out-of-sample prediction is done by running the sampling step for 500 iterations with each consecutive iteration reading in and conditioning on the next value of $\beta$ and $\boldsymbol{b}$ found during the training period. The first 200 iterations are used for burn-in. The remaining 300 iterations are thinned by discarding every other observation, leaving a final predictive sample of 150 observations.

>>>>

This paper adds another layer of complexity by assuming a heterogeneous set of experts, most of whom only make one or two probability assessments over the hundred or so days before the event resolves. This means that the decision-maker faces a different group of experts  every day, with only a few experts returning later on for a second round of advice. The problem at hand is therefore strikingly different from many time-series estimation problems, where one has an observation at every time point -- or almost every time point. As a result, standard time-series procedures like ARIMA (see, e.g., \citet{mills1991time}) are not directly applicable. This paper introduces a time-series model that incorporates self-reported expertise and captures a sharp and well-calibrated estimate of the crowd belief. The model is highly interpretable and can be used for:
\begin{itemize}
\item borrowing strength across the hierarchy to analyze under- and overconfidence in different groups of experts,
%\item confidence bands for the crowd belief,
\item accurate probability forecasts, and
\item many question-specific quantities that have easy interpretations, such as expert disagreement and problem difficulty, and can be used to gain novel insight in the social sciences. 
\end{itemize}

This paper begins by describing our geopolitical database. The paper then introduces a dynamic hierarchical model for capturing the crowd belief. The model is estimated in a two-step procedure: first, a sampling step produces constrained parameter estimates via Gibbs sampling (see \citet{geman1984stochastic} for the original introduction of Gibbs sampling); second, a calibration step transforms these estimates to their unconstrained equivalents via a one-dimension optimization procedure. The model introduction is followed by the first evaluation section that uses synthetic data to study how accurately the two-step procedure can estimate parameter values. The second evaluation section applies the model to real-world forecasting data. The paper concludes with a discussion on future research directions and model limitations.

@@

The experts made updates on a very infrequent basis: the average number of forecasts made by a single expert in one day was around 0.017, and the average group-level response rate was around 13.5 forecasts per day. Given that the group of experts is large and diverse, the resulting dataset is very sparse. Tables \ref{DataStats} and \ref{ExpertiseTable} provide relevant summary statistics on the data. Notice that the distribution of the self-reported expertise is skewed to the right and that some questions remained active longer than others.  For more details on the dataset and its collection see \citet{ungar2012good}.

@@

\subsection{Sampling Step}
\label{sampling_step}
Given that $\left(a \boldsymbol{b}, X_{t,k}/a, a^2 \tau_k^2\right) \neq \left(\boldsymbol{b}, X_{t,k}, \tau_k^2\right)$ for any $a > 0$ yield the same  likelihood for $\boldsymbol{Y}_{t,k}$, the model as described by Equations (\ref{observedpr}) and (\ref{hiddenpr}) is not identifiable. As a result, the parameter estimates tend to drift during the sampling process. A well-known solution is to choose one of the elements of $\boldsymbol{b}$, say $b_3$, as the reference point and fix $b_3 = 1$.  In Section \ref{syntheticData} we provide a guideline for choosing the reference point. Denote the constrained version of the model by
\begin{eqnarray*}
\boldsymbol{Y}_{t, k} &=&  \boldsymbol{M}_k \boldsymbol{b} (1) X_{t, k}(1)+ \boldsymbol{v}_{t, k}\\
X_{t, k}(1) &=& \gamma_k(1) X_{t-1, k}(1) + w_{t, k}\\
\boldsymbol{v}_{t, k} | \sigma^2_k(1) &\stackrel{i.i.d.}{\sim}& \mathcal{N}_{I_k}(\boldsymbol{0}, \sigma^2_k(1) \boldsymbol{I}_{I_k})\\
w_{t, k} | \tau^2_k(1) &\stackrel{i.i.d.}{\sim}& \mathcal{N}\left(0, \tau^2_k(1)\right),
\end{eqnarray*}
where the trailing input parameter emphasizes the constraint $b_{3} = 1$. Given that this version is identifiable, estimates of the model parameters can be obtained. Denote the estimates by placing a hat on the parameter symbol. For instance, $\hat{\boldsymbol{b}}(1)$ and $\hat{X}_{t, k}(1)$ represent the estimates of $\boldsymbol{b}(1)$ and $X_{t, k}(1)$, respectively. These estimates are found via Gibbs sampling that only makes use of standard distributions. See Appendix \ref{appendix} for the technical details of the sampling step, and, e.g., \citet{gelman2003bayesian} for a discussion on the general principles of Gibbs sampling.

\subsection{Calibration Step}
\label{calibration_step}
Given that the model parameters can be estimated by fixing $b_3$ to any constant, the next step is to search for the constant that gives an optimally sharp and calibrated estimate of the hidden process.  This section introduces an efficient procedure that finds the optimal constraint without requiring any additional runs of the sampling step. First, assume that parameter estimates  $\hat{\boldsymbol{b}}(1)$ and $\hat{X}_{t, k}(1)$ have already been obtained via the constrained sampling step described in Section \ref{sampling_step}. Given that for any $\beta \in \R / \{0\}$,
\begin{eqnarray*}
\boldsymbol{Y}_{t, k} &=&  \boldsymbol{M}_k \boldsymbol{b} (1) X_{t, k}(1)+ \boldsymbol{v}_{t, k}\\
 &=&   \boldsymbol{M}_k \left(\boldsymbol{b} (1) \beta \right) \left(X_{t, k}(1)/\beta\right)  + \boldsymbol{v}_{t, k}\\
&=&   \boldsymbol{M}_k \boldsymbol{b} (\beta) X_{t, k}(\beta) + \boldsymbol{v}_{t, k},
\end{eqnarray*}
the parameter values under $b_{3} = \beta$ can be obtained from $\boldsymbol{b} (\beta) = \boldsymbol{b} (1) \beta $ and $X_{t, k}(\beta) = X_{t, k}(1)/\beta$. 
%Furthermore, an estimate of $X_{t, k}(\beta)$ is given by $\hat{X}_{t, k}(1)/\beta$. 
This means that $X_{t,k} = X_{t,k}(1)/\beta$ when $\beta$ is equal to the true value of $b_3$. Given that the hidden process $X_{t, k}$ is assumed to be sharp and well-calibrated, $b_3$ can be estimated with the value of $\beta$ that simultaneously maximizes the sharpness and calibration of $\hat{X}_{t, k}(1) / \beta$. A natural criterion for this maximization is given by the class of \textit{proper scoring rules} that combine sharpness and calibration (\citet{gneiting2008rejoinder, buja2005loss}). Due to the possibility of \textit{complete separation} in any one question (see, e.g., \citet{gelman2008weakly}), the maximization must be performed over multiple questions. Therefore,
\begin{eqnarray}
 \hat{\beta}&=&  \argmax_{\beta \in \R / \{0\}} \sum_{k=1}^K \sum_{t=1}^{T_k}  S\left(Z_k, \hat{X}_{k,t}(1) / \beta \right) \label{OSE}
\end{eqnarray}
where $Z_k \in \{0, 1\}$ indicate whether the event associated with the $k$th question happened $\left(Z_k = 1\right)$ or did not happen $\left(Z_k = 0\right)$. The function $S$ is a strictly proper scoring rule such as the negative Brier score (\citet{Brier})
\begin{eqnarray*}
S_{BRI}(Z, X) &=& -(Z - \logit^{-1}(X))^2
\end{eqnarray*}
or the logarithmic score (\citet{good1952rational})
\begin{eqnarray*}
S_{LOG}(Z, X) &=& Z \log\left(\logit^{-1}(X)\right) + (1-Z) \log\left(1-\logit^{-1}(X)\right)
\end{eqnarray*}
Given that it is not clear which rule should be used for predicting geopolitical events, the SAC-procedure is evaluated  separately under both rules in Sections \ref{syntheticData} and \ref{realData}. Once $\hat{\beta}$ has been computed, estimates of the unconstrained model parameters are given by
\begin{eqnarray}
 \hat{X}_{t,k}&=& \hat{X}_{k,t}(1) / \hat{\beta} \nonumber\\
 \hat{\boldsymbol{b}}&=& \hat{\boldsymbol{b}}(1) \hat{\beta} \nonumber\\
   \hat{\tau}_{k}^2&=& \hat{\tau}_{k}^2(1)/ \hat{\beta}^2\nonumber\\
  \hat{\sigma}_{k}^2&=& \hat{\sigma}_{k}^2(1)\nonumber\\
  \hat{\gamma}_{k}&=& \hat{\gamma}_{k}(1)\nonumber
\end{eqnarray}
Notice that estimates of $\sigma^2_k$ and $\gamma_k$ are not affected by the constraint. Therefore their constrained and unconstrained versions are the same.

@@

If the future event can take upon $M > 2$ possible outcomes, the hidden state $X_{t,k}$ is extended to a vector of size $M-1$ and one of the outcomes, e.g., the $M$th one, is chosen as the base-case to ensure that the probabilities will sum to one at any given time point. Each of the remaining  $M-1$ possible outcomes is represented by an observed process similar to Equation (\ref{observedpr}). Given that this multinomial extension is equivalent to having $M-1$ independent binary-outcome models, the estimation and properties of the model are easily extended to the multi-outcome case. This paper focuses on binary-outcomes because it is the simplest and most commonly encountered setting in practice.

@@

The goal in this section is to evaluate the ability of the \textit{Sample-And-Calibrate}-procedure to capture true parameter values. The synthetic data are not generated directly from the model for two reasons: (i) showing good performance on a dataset directly generated from the model assumptions is hardly news, and (ii) it is unclear how the dynamic part of the model can be used to produce well-calibrated hidden states. The latter is important for our interpretation of the hidden process as a sharp and well-calibrated version of the crowd belief.

To ensure that the hidden process for question $k$ is well-calibrated, we generate a path of standard Brownian motion until time $T_k$. If $Z_{t,k}$ denotes the value of the path at time $t$, then
\begin{eqnarray*}
Z_k &=& \mathbbm{1}\left( Z_{T_k, k}  > 0\right)\\
 X_{t,k} &=& \logit \left[ \Phi \left(\frac{Z_{t, k}}{\sqrt{T_k- t}} \right) \right]
\end{eqnarray*}
gives a sequence of $T_k$ calibrated logit-probabilities for the event $Z_k = 1$. Using this procedure we generate the hidden process for $K$ questions, each with a time horizon of $T_k = 101$. The questions involve 50 experts allocated evenly among five expertise groups. Each expert gives one probability forecast per day with the exception of time $t = 101$ when the event resolves. These are generated by applying bias and noise to the hidden process $X_{t,k}$ for $t = 1, \dots, 100$ (see Equation (\ref{observedpr})). Our simulation study considers a three-dimensional grid of parameter values:
\begin{eqnarray*}
\sigma^2 &\in& \{1/2, 1, 3/2, 2, 5/2\}\\
\beta &\in& \{1/2, 3/4, 1, 4/3, 2/1\}\\
K &\in& \{20, 40, 60, 80, 100\},
\end{eqnarray*}
where $\beta$ gives the bias vector by letting $\boldsymbol{b} = [1/2, 3/4, 1, 4/3, 2/1]^T \beta$. Forty synthetic datasets are generated for each combination of $\sigma^2$, $\beta$, and $K$ values. The SAC-procedure runs for 200 iterations of which the first 100 are used for burn-in. The accuracy of the estimated hidden process and bias vector are measured with the average quadratic loss in the probability space, $\sum_{k=1}^K \sum_{t=1}^{100} ( \logit^{-1}(\hat{X}_{t,k}) - \logit^{-1}(X_{t,k}))^2 / (100K)$, and the average quadratic loss, $|| \boldsymbol{b} - \hat{\boldsymbol{b}}||^2/5$, respectively.

@@

\begin{figure}[h!]
\vspace{-5.5em}
\centering
 \hspace{-25em}	\includegraphics{Figures/LegendSynthetic} % requires the graphicx package

\vspace{0em}
\hspace{2.5em}Hidden Process
\vspace{-0.5em}

@@

\vspace{0.5em}
\hspace{2.5em}Bias Vector
\vspace{-0.5em}

@@

\caption[Optional caption for list of figures]{Comparing SAC that optimizes over the logarithmic score ($\text{SAC}_{\text{LOG}}$), SAC that optimizes over the Brier score ($\text{SAC}_{\text{BRI}}$), and the Exponentially Weighted Moving Average (EWMA) under synthetic data. The top row presents the accuracy to capture the hidden process. The bottom row presents the accuracy to capture the bias vector. }
\label{Synthetic}
\end{figure}

Figure \ref{Synthetic} summarizes the results in a series of plots. The first row presents the marginal effects of the three grid variables on the accuracy to which SAC and the Exponentially Weighted Moving Average (EWMA) (see Section \ref{forecasting} for more details on EWMA) estimate the hidden process. EWMA serves as a baseline for comparisons. Unfortunately, EWMA does not produce an estimate of the bias vector. Therefore it cannot be used as a baseline in the second row that measures the accuracy in estimating the bias vector.  Given that the plots were computed by choosing one grid variable (e.g., $\beta$) at a time and averaging over the remaining two variables  (e.g., $K$ and $\sigma^2$), each value in the plots represents an average of a $40 \times 5^2 = 1,000$ values.

Both $\text{SAC}_{\text{LOG}}$ and  $\text{SAC}_{\text{BRI}}$ estimate the hidden process and the bias vector very accurately. Apart from $\text{SAC}_{\text{LOG}}$ outperforming $\text{SAC}_{\text{BRI}}$ uniformly across all parameter values, the two approaches behave very similarly.  They suffer  from the increasing level of noise in the expert logit-probabilities, estimate the bias vector less accurately when $\beta$ is large, and, unlike EWMA, make greater use of the increasing number of questions.  EWMA is able to outperform SAC only when the bias among the experts or the number of problems is very small.

Interestingly, SAC estimates the hidden process more accurately when the experts are over-confident (large $\beta$) rather under-confident (small $\beta$). To understand this result, assume that the experts in the third group are highly under-confident. Therefore their logit-probabilities are expected to be closer to zero than the corresponding hidden states. After adding white noise to these expected logit-probabilities, they are likely to cross to the other side of zero. If the sampling step fixes $b_3 = 1$, as it does in our case, the third group is treated unbiased and some of the constrained estimates of the hidden states are likely to be on the other side of zero as well. Unfortunately, this discrepancy cannot be corrected by the calibration step that is restricted to shifting the constrained estimates either closer or further away from zero but not across it. To maximize the likelihood of having all the constrained estimates on the right side of zero and hence avoiding the discrepancy, the reference point in the sampling step should be chosen with care. A helpful guideline is to fix the element of $\boldsymbol{b}$ that is apriori believed  to be the largest.

@@

\section{Geopolitical Data Results}
\label{realData}
\noindent
This section presents results for the real-world data described in Section \ref{data}. The goal is to provide application specific insight by discussing the specific research objectives itemized in Section \ref{intro} and also to evaluate SAC in terms of predictive power and calibration. Before presenting the results, however, we discuss two practical matters that must be taken into account when aggregating real-world  probability forecasts.

@@

The second matter is related to the distribution of the class labels in the data. If the set of occurrences is much larger than the set of non-occurrences (or \textit{vice versa}), the dataset is called \textit{imbalanced}. On such data the modeling procedure can end up over-focusing on the larger class, and as a result, give very accurate forecast performance over the larger class at the cost of performing poorly over the smaller class (see, e.g., \citet{chen2009learning, wallace2012class}). Fortunately, it is often possible to use a well-balanced version of the data. The first step is to find a partition $S_0$ and $S_1$ of the question indices $\{1, 2, \dots, K\}$ such that the equality $\sum_{k \in S_0} T_k  = \sum_{k \in S_1} T_k$ is as closely approximated as possible. This is equivalent to an NP-hard problem known in computer science as the \textit{Partition Problem}:  determine whether a given set of positive integers can be partitioned into two sets such that the sums of the two sets equal to each other (see, e.g., \citet{karmarkar1982differencing, hayes2002easiest}). A simple solution is to use a greedy algorithm that iterates through the values of $T_k$ in descending order, assigning each $T_k$ to the subset that currently has the smaller sum (see, e.g. \citet{kellerer2004knapsack, gent1996phase} for more details on the \textit{Partition Problem}). After finding a well-balanced partition, the next step is to assign the class labels such that the labels for the questions in $S_x$ are equal to $x$ for $x = 0$ or $1$. Recall from section \ref{calibration_step} that $Z_k$ represents the event indicator associated with the $k$th question. To define a balanced set of indicators $ \tilde{Z}_k$ for all $k \in S_x$, let
\begin{align*}
 \tilde{Z}_k &= x\\
\tilde{p}_{i,t,k} &=  \begin{cases} 
1-p_{i,t,k}, & \text{if } Z_k = 1-x,\\
p_{i,t,k}, & \text{if } Z_k = x,
\end{cases}
\end{align*}
where $i = 1, \dots, I_k$, and $t = 1, \dots, T_k$. The resulting set $$\left\{\left(\tilde{Z}_k, \{\tilde{p}_{i,t,k} | i = 1, \dots, I_k, t = 1, \dots, T_k \}\right) \right\}_{k=1}^K$$ is a balanced version of the data. 
This procedure was used to balance our real-world dataset both in terms of events and time points. The final output splits the events exactly in half ($|S_0| = |S_1| = 83$) such that number of time points in the first and second halves are 8,737 and 8,738, respectively.

\subsection{Out-of-Sample Forecasting}
\label{forecasting}
This section is motivated by decision-making. 
%For instance, recall our example with an investor who consults a group of experts on a daily basis in order to assess whether a stock index will end up trading above some threshold on a given date. 
The goal is to evaluate the out-of-sample predictive performance of SAC against several other probability aggregation procedures. The models are allowed to utilize a training set before making predictions on an independent testing set. To clarify some of the upcoming notation, let $S_{train}$ and $S_{test}$ be index sets that partition the data into training and testing sets of sizes $|S_{train}| = N_{train}$ and $|S_{test}| = 166 - N_{test}$, respectively. This means that the $k$th question is in the training set if and only if $k \in S_{train}$. The competing models are as follows:

@@

\item \textit{The Sample-And-Calibrate procedure both under the Brier ($\text{SAC}_{\text{BRI}}$) and the logarithmic score ($\text{SAC}_{\text{LOG}}$)}. The model is first fit on the training set by running the sampling step for 3,000 iterations of which the first 500 iterations are used for burn-in. The remaining 2,500 observations are thinned by keeping every fifth observation. The calibration step is performed for the final 500 observations. The out-of-sample prediction is done by running the sampling step for 500 iterations with each consecutive iteration reading in and conditioning on the next value of $\beta$ and $\boldsymbol{b}$ found during the training period. The first 200 iterations are used for burn-in. The remaining 300 iterations are thinned by discarding every other observation, leaving a final predictive sample of 150 observations.
